Colonies - Compute Continuums across Platforms

### 2023-04-01 paper_2023/paper.tex
wip

In addition to a technical description of the Colonies framework, the paper also describes some potential use cases. The paper describe how Colonies can be leveraged to build a scalable remote sensing platform on Kubernetes, serve as a building block for edge computing, and how it can be integrated with HPC platforms. Finally, the paper presents a performance investigation, as well as a scalability and robustness evaluation.

@@

The next section presents related work. The next section describes related work. Section X ...

@@

\begin{figure}
     \centering
     \begin{subfigure}[b]{0.3\textwidth}
         \centering
         \includegraphics[scale=0.39]{arch.png}
         \caption{Process assignment steps.}
         \label{fig:y equals x}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.3\textwidth}
         \centering
         \includegraphics[scale=0.39]{process.png}
         \caption{Manipulation of a process.}
         \label{fig:three sin x}
     \end{subfigure}
     \caption{Process management via Colonies HTTP API. Note that only the assigned executor has write access to the process database entry.}
     \label{fig:three graphs}
\end{figure}

@@

The Colonies server acts as a job broker for executors, almost like an employment agency. When an executor connects to the Colonies server, the server hang the incomming HTTP connection\footnote{An alternative protocol is to use WebSockets or gPRC to communicate with the Colonies server.} until the executor is assigned a process, or until a connection timer expires. Note that the Colonies server does not connect to any executor. Rather, it is the responsibility of the executors to connect to the Colonies server, enabling them to be deployed anywhere on the Internet, even behind firewalls or commercial telco networks.

Figure \ref{fig:fig1} illustrates an executor submitting a function specification that is later assigned to another executor. Executors have to specify to the Colonies server which functions they are capable of executing. The Colonies server makes sure that the conditions (or requirmenets) of a function specification matches the capability of the executors.

\begin{figure}[h]
	\centering
    \includegraphics[scale=0.35]{function_spec.png}
	\caption{Example of a function spec.}
	\label{fig:fig2}
\end{figure}

Figure \ref{fig:fig2} shows an example of a function specification. The \emph{maxexectime} attribute specifies the maximum time an executor may run a process (in this case, 100 seconds). Before a process is assigned to an executor, the Colonies server updates the process entry in the process table database and calculates a deadline for when the process must be finished. The server then regularly checks for any running processes that have exceeded their deadlines. If such process is detected, it is reset, allowing it to be re-assigned to other executors. This scheme is completely stateless, so if the Colonies server crashes, another Colonies server can take over and look for processes that have exceeded their deadline. Moreover, if an executor assigned to a process crashes, the process is simply re-executed by another executor.

To scale up a system, more executors can be deployed. Scaling down, however, can be more challenging. One solution is to select a set of executors to be removed and then starv them out by denying them new process assignments. Another, simpler solution, is to immediately destroy the executors, or if possible notify them in advance so that they can gracefully unregister.

The \emph{maxexectime} failsafe mechanism ensures that processes will eventually be executed even in the case of failures. This mechanism also relieves the burden of user to check if a process has been executed or not, as they can simply look up the process in the database to get its current status. It is also possible to subscribe on process changhes using an event-driven protocol, but this may be impractical if a process runs for a very long time (e.g. days or weeks).

\begin{figure}[h]
	\centering
    \includegraphics[scale=0.4]{raft.png}
	\caption{cron management}
	\label{fig:fig2}
\end{figure}

\subsubsection{Concurreny and synchronization}
Synchronization is esstential to prevent data inconsistency and race conditions when accessing shared resources concurrently with multiple threads. However, synchronization also slows down execution as only one thread can access critical sections at a time. By carefully designing multithreaded applications and employing the right synchronization techniques, it possible to minimize the performance impact while still ensuring data consistency and correctness.

The purpose of all Colonies APIs is to either modify the database or read information from the database. A request

Colonies is designed to be stateless, meaning that the Colonies server does not keep any information about between API requests. In other words, each request is handled independently, without relying on any information from previous requests. Once the database has been updated,

if an instance fails or becomes unavailable, subsequent requests can be easily redirected to another

. After a The Colonies server ensure that only the assigned executor has write access to the process entry in the database.

Colonies ensures that exactly one executor is assigned to a specific process, so only the assigned executor has permission to set output values and close the process as successful. However, there is a window of time during which an executor may execute a process.

\subsubsection{Workflows}
TODO
\begin{figure}[h]
	\centering
    \includegraphics[scale=0.32]{workflow.png}
	\caption{cron management}
	\label{fig:fig3}
\end{figure}

@@

dt = -1000000000 * 60 * 60 * 24
process.PriorityTime = int64(process.FunctionSpec.Priority)*dt + submissionTime.UnixNano()

@@

\section{Evaluation}
\subsection{Implementation}
\begin{lstlisting}[language=c]
gen_nums = Function(gen_data, colonyid, executortype="edge")
square1 = Function(square, colonyid, executortype="cloud")
square2 = Function(square, colonyid, executortype="cloud")
sum = Function(square, colonyid, executortype="browser")

wf = ColoniesWorkflow("localhost", 50080, colonyid, executor_prvkey)
wf >> gennums
gennums >> square1
gennums >> square2
[square1, square2] >> sum
res = wf.execute()
\end{lstlisting}

\subsection{References}
TODO
\bibliographystyle{unsrtnat}
\bibliography{references}  %%% Uncomment this line and comment out the ``thebibliography'' section below to use the external .bib file (using bibtex) .

%%% Uncomment this section and comment out the \bibliography{references} line above to use inline references.
% \begin{thebibliography}{1}

% 	\bibitem{kour2014real}
% 	George Kour and Raid Saabne.
% 	\newblock Real-time segmentation of on-line handwritten arabic script.
% 	\newblock In {\em Frontiers in Handwriting Recognition (ICFHR), 2014 14th
% 			International Conference on}, pages 417--422. IEEE, 2014.

% 	\bibitem{kour2014fast}
% 	George Kour and Raid Saabne.
% 	\newblock Fast classification of handwritten on-line arabic characters.
% 	\newblock In {\em Soft Computing and Pattern Recognition (SoCPaR), 2014 6th
% 			International Conference of}, pages 312--318. IEEE, 2014.

% 	\bibitem{hadash2018estimate}
% 	Guy Hadash, Einat Kermany, Boaz Carmeli, Ofer Lavi, George Kour, and Alon
% 	Jacovi.
% 	\newblock Estimate and replace: A novel approach to integrating deep neural
% 	networks with existing applications.
% 	\newblock {\em arXiv preprint arXiv:1804.09028}, 2018.

% \end{thebibliography}

>>>>

    In addition to a technical description of the Colonies framework, the paper also describes some potential use cases. The paper describe how Colonies can be leveraged to build a remote sensing platform on Kubernetes, serve as a building block for edge computing, implement a serverless FaaS (Function-as-a-Service), and how it can be integrated with HPC platforms. Finally, the paper presents a performance investigation, as well as  scalability and robustness evaluation.

@@

@@

\begin{equation}
    priority_{time}=priority \cdot -10^9 \cdot 60 \cdot 60 \cdot 24 + submission_{time}  
\end{equation}

\begin{figure}
     \centering
     \begin{subfigure}[b]{0.3\textwidth}
         \centering
         \includegraphics[scale=0.37]{arch.png}
         \caption{Process assignment steps.}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.3\textwidth}
         \centering
         \includegraphics[scale=0.37]{process.png}
         \caption{Manipulation of a process.}
     \end{subfigure}
     \caption{Process management via Colonies HTTP API. Note that only the assigned executor has write access to the process database entry.}
     \label{fig:process_assignment}
\end{figure}

@@

The Colonies server acts as a job broker for executors, almost like an employment agency for people. When an executor connects to the Colonies server, the server hang the incomming HTTP connection\footnote{An alternative protocol is to use WebSockets or gPRC to communicate with the Colonies server.} until the executor is assigned a process, or until a connection timer expires. Note that the Colonies server does not connect to any executor. Rather, it is the responsibility of the executors to connect to the Colonies server. This enables executors to be deployed anywhere on the Internet, even behind firewalls, commercial telco networks, or in application running in web browsers.

Figure \ref{fig:process_assignment} illustrates an executor submitting a function specification that is later assigned to another executor. Executors have to specify to the Colonies server which functions they are capable of executing. The Colonies server makes sure that the conditions (i.e requirmenets) of a function specification matches the capability of an executor.

\begin{figure}[h]
	\centering
    \includegraphics[scale=0.35]{function_spec.png}
	\caption{Example of a function spec.}
	\label{fig:function_spec}
\end{figure}

Figure \ref{fig:function_spec} shows an example of a function specification. The \emph{maxexectime} attribute specifies the maximum time an executor may run a process (in this case, 100 seconds). Before a process is assigned to an executor, the Colonies server updates the process entry in the process table database and calculates a deadline for when the process must be finished. The server then regularly checks for any running processes that have exceeded their deadlines. If such process is detected, it is reset, allowing it to be re-assigned to other executors. This scheme is completely stateless, so if the Colonies server crashes, another Colonies server can take over and look for processes that have exceeded their deadline. Moreover, if an executor assigned to a process crashes, the process is simply re-executed by another executor.

Making it possible to specify maximum execution time is a simple but powerful mechanism. To scale up a system, more executors can simply be deployed. Scaling down, however, can be more challenging. One solution is to select a set of executors to be removed and then starv them out by denying them new process assignments. Another, simpler solution, is to immediately destroy the executors, or if possible notify them in advance so that they can gracefully unregister.

The \emph{maxexectime} failsafe mechanism ensures that processes will eventually be executed even in the case of failures. This mechanism also relieves the burden of user to check if a process has been executed or not, as they can simply look up the process in the database to get its current status. The Colonies framework also supports subscribing on process changes using an event-driven protocol, but this may be impractical if a process runs for a very long time (e.g. days or weeks).

Utilizing the \emph{maxexectime} failsafe mechanism not only enhances system reliability, but also provides an opportunity to apply of Chaos engineering \cite{chaos_engineering}. For example, deliberately terminating executors using a Chaos monkey. If executors are deployed on Kubernetes, Kubernetes will automatically redeploy terminated executors. The constant replacement of executor instances ensures that the system is capable of gracefully tolerating failures.

\begin{figure}[h]
	\centering
    \includegraphics[scale=0.5]{raft.png}
	\caption{High-availability deployment of Colonies servcer.}
	\label{fig:ha_deployment}
\end{figure}

\subsubsection{Concurreny and synchronization}
Synchronization is esstential to prevent data inconsistency and race conditions when accessing shared resources concurrently with multiple threads. However, synchronization can also slows down execution as only one thread can access critical sections at a time. By carefully designing multithreaded applications and employing the right synchronization techniques, it possible to minimize the performance impact while still ensuring data consistency and correctness.

The primary objective of all Colonies API requests is to either alter a state stored in the database or retrieve information from it. The Colonies framework is designed to be stateless, meaning that the Colonies server does not keep any information between requests. In other words, each request is handled independently, without relying on any information from previous requests.

One example of a request is the assign request, which allocates a waiting process from the database to an executor. Given the multi-threaded nature of the Colonies server, it is essential that the assign request is synchronized to ensure that only one thread at a time can modify the database and update the process table, thus preventing multiple executors from being assigned to the same process. As only the assigned executor can modify a running process and only one executor must be assigned to a process, synchronization is required for the assign request. It is worth noting that synchronization is not necessary for other requests, such as the submit request or the close request. As the submit request only add new entries to the process table, there is no race conditions. The close request set the output of the function innovation and updates the process state to either successful or failed in the process table. Since there exists only one assigned executor there is no need for synchronization.

To minimize downtime in the event of hardware or software failures, there is a need for high-availability and make use of a cluster of Colonies servers. If one Coloniser server crashes, an executor simply need to re-send a failed request, which will then be served by another Colonies server replica. However, by introducing multiple Colonies servers, there is again a risk of race conditions when assigning processes to executors. This means that the Colonies server replicas must coordintate which replica server incomming assign requests so that precisely one executor is assigned to a process.

Raft \cite{raft} is a consensus algorithm specifically designed to manage a replicated log within a distributed system. It functions within a cluster of servers, where a single server takes on the role of leader while the remaining servers act as followers. The leader is responsible for managing the replicated log, processing client requests, and replicating entries to the followers. Followers passively replicate the leader's log and participate in leader elections. The leadership can change over time due to elections triggered by timeouts or other factors.

Incorporating Raft into the Colonies framework allows incoming assign requests to be directed towards the leading Colonies server, ensuring that only one Colonies server replica handles these requests. Figure \ref{fig:ha_deployment} shows an overview of a high-availability Colonies deployment. A new leader is elected in the event that a Colonies server replica fails. The Raft protocol also enables seamless updates to the Colonies server software by making it possible to upgrade each replica individually. As a result, Colonies becomes well-suited for Kubernetes deployments.

\subsubsection{Workflows}
TODO
\begin{figure}[h]
	\centering
    \includegraphics[scale=0.30]{workflow.png}
	\caption{cron management}
	\label{fig:fig3}
\end{figure}

@@

@@

\section{Evaluation}
\subsection{Implementation}

\bibliographystyle{unsrt}
\bibliography{references}
